\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-04-15 archive.tex
Update archive.tex
libraries to language figure starting

\begin{figure}
  \includegraphics[width = \textwidth]{figures/dpl/list}
  \includegraphics[width = \textwidth]{figures/DeepLogo_4_11}
  \caption{}\label{exampleDSL}
\end{figure}

@@

Rather than start with a domain-specific basis and then enrich it over successive wake/sleep cycles -- as these prior experiments have done --
could we instead start with a highly generic basis,
and then learn a library encoding something closer to the basic domain language?
We first consider learning a basic language for equations describing physical laws,
starting \emph{without} concepts like vectors or forces
and instead providing sequences of real numbers and Lisp-style sequence manipulation.
We collected 57 equations for physical laws from AP and MCAT physics `cheat sheets' --
forms like Coulomb's law and expressions for the center of mass of a set of objects --
along with
reference vector algebra identities from machine learning textbooks,
and tasked our agent with inferring programs calculating these input/output mappings.
In solving these tasks, \system first discovers the building blocks of vector algebra,
like inner products, vector sums, and norms,
and then uses this vocabulary to infer library routines for common
physical relationships like inverse square laws and integrating second derivatives over time.

@@

    \begin{tabular}{ll}      
      \begin{tabular}{l}
        \code{[1\, 9]}$\to $\code{2}\\
        \code{[5\, 3\, 8]$\to$3}\\
        \blueCode{$f(\ell) = $($f_5$ $\ell$)}
      \end{tabular}&
      \begin{tabular}{l}
        \code{[[2 1]\, []]}$\to $\code{[1 0]}\\
        \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
        \blueCode{$f(\ell) = $($f_2$ $f_4$ $\ell$)}
      \end{tabular}\\\\

@@

      \begin{tabular}{l}
        \code{0, [9 2 3]$\to $9}\\
        \code{3, [0 2 8 4 5 6]$\to $4}\\
        \blueCode{$f($n$,$l$) = $($f_{10}$ l n)}
      \end{tabular}&
      \begin{tabular}{l}
        \code{1, [9 2 3]$\to $9}\\
        \code{4, [0 2 8 1 5 6]$\to $1}\\
        \blueCode{$f($n$,$l$) = $($f_{10}$ l (+ 1 n))}
      \end{tabular}\\\\
      \begin{tabular}{l}
        \code{3$\to $[-3 -2 -1]}\\
        \code{4$\to $[-4 -3 -2 -1]}\\
        \blueCode{$f($n$) = $($f_{8}$ 0 n)}
      \end{tabular}&
      \begin{tabular}{l}
        \code{2$\to $[2 1]}\\
        \code{4$\to $[4 3 2 1]}\\
        \blueCode{$f($n$) = $($f_{8}$ 0 (- 0 n))}
      \end{tabular}      
    \end{tabular}
    \\\midrule 
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \begin{tabular}{ll}
      \begin{tabular}{l}
        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
          \hspace{\helpSize}($f_0$: \emph{fold})
      \end{tabular}&
      \begin{tabular}{l}
        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
          \hspace{\helpSize}($f_1$: \emph{unfold})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_2$: \emph{map})
      \end{tabular}&
      \begin{tabular}{l}
                  \greenCode{$f_3($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (if (f x) a (cons x a))))}\\
          \hspace{\helpSize}($f_3$: \emph{filter})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
        \hspace{\helpSize}($f_4$: \emph{generalization of range})
      \end{tabular}&
      \begin{tabular}{l}
        \greenCode{$f_5($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
        \hspace{\helpSize}($f_5$: \emph{length})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($n$,$p$,$l$)\,=\,$($f_1$ p n car l)}\\
        \hspace{\helpSize}($f_6$: \emph{specialization of unfold})
      \end{tabular}&
      \begin{tabular}{l}
        \greenCode{$f_7($p$)\,=\,$($f_4$ ($\lambda$ (x) x) p 0)}\\
        \hspace{\helpSize}($f_7$: \emph{another generalization of range})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_8($n$,$m$)\,=\,$($f_4$ ($\lambda$ (x) (- n x)) (eq? 0) m)}\\
        \hspace{\helpSize}($f_7$: \emph{count downwards})
      \end{tabular}&
      \begin{tabular}{l}
        \greenCode{$f_9($n$)\,=\,$($f_7$ (eq? n))}\\
        \hspace{\helpSize}($f_9$: \emph{range})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_{10}($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_9$ n) l))}\\
        \hspace{\helpSize}($f_{10}$: \emph{index})
      \end{tabular}&

>>>>

\begin{figure}
  \includegraphics[width = 1.2\textwidth]{figures/DeepLogo_4_11}
  \includegraphics[width = \textwidth]{figures/dpl/list}
  \caption{}\label{exampleDSL}
\end{figure}

@@

Rather than start with a domain-specific basis and then enrich it over successive wake/sleep cycles -- as these prior experiments have done --
could we instead start with a highly generic basis,
and then learn a library encoding something closer to the basic domain language?
We first consider learning a basic language for equations describing physical laws,
starting \emph{without} concepts like vectors or forces
and instead providing Lisp-style sequence manipulation primitives like
\code{map} and \code{fold}.
We collected 57 equations for physical laws from AP and MCAT physics `cheat sheets' --
forms like Coulomb's law and expressions for the center of mass of a set of objects --
along with
reference vector algebra identities from machine learning textbooks,
and tasked our agent with inferring programs calculating these input/output mappings.
In solving these tasks, \system first discovers the building blocks of vector algebra,
like inner products, vector sums, and norms,
and then uses this vocabulary to infer library routines for common
physical relationships like inverse square laws and integrating second derivatives over time.

\begin{figure}
  \centering
  \begin{tabular}{lll}\toprule
    &\multicolumn{1}{c}{Physical Laws}&\multicolumn{1}{c}{Recursive Functions}\\\midrule
    \normalsize \rotatebox[origin=c]{90}{\pop{Programs} \& Tasks}&
    % physical laws programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Center-of-mass:\\
      \end{tabular}\\\\
      \begin{tabular}{l}
        Coulomb's law:\\
      \end{tabular}\\\\
      \begin{tabular}{l}
        Integrating position over time:\\
        \end{tabular}
      \end{tabular}
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Duplicate each list element:\\
        \code{[9 2]$\to $[9 9 2 2]}\\
        \code{[5 0 3]$\to $[5 5 0 0 3 3]}\\
        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
        \blueCode{$f(l) = $($f_0$ ($\lambda$ (x a) (cons x (cons x a))) l nil)}        
      \end{tabular}\\\\

      \begin{tabular}{l}
        Lengths of list-of-lists:\\
        \code{[[2 1] [1 2 9]\, []]}$\to $\code{[2 3 0]}\\
        \code{[[5 2 9 3 7 5 2]]}$\to $\code{[7]}\\
        \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
        \blueCode{$f(\ell) = $($f_2$ ($\lambda$ (z) ($f_4$ z)) $\ell$)}
      \end{tabular}\\\\
      \begin{tabular}{l}
        1-index into list:\\
        \code{1, [9 2 3]$\to $9}\\
        \code{2, [1 5 8]$\to $5}\\
        \code{4, [0 2 8 1 5 6]$\to $1}\\
        \blueCode{$f($n$,$l$) = $($f_7$ l (+ 1 n))}
      \end{tabular}\\\\

      \end{tabular}
    \\\midrule
    \rotatebox[origin=c]{90}{\popp{Library}}&
    % physical laws library
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
          \hspace{\helpSize}($f_0$: \emph{fold})
      \end{tabular}\\
      \begin{tabular}{l}
        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
          \hspace{\helpSize}($f_1$: \emph{unfold})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_2$: \emph{map})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_3($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
        \hspace{\helpSize}($f_3$: \emph{count upward until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
        \hspace{\helpSize}($f_4$: \emph{length})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_5($p$)\,=\,$($f_3$ ($\lambda$ (x) x) p 0)}\\
        \hspace{\helpSize}($f_5$: \emph{count upward from 0 until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($n$)\,=\,$($f_5$ (eq? n))}\\
        \hspace{\helpSize}($f_6$: \emph{range})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_7($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_6$ n) l))}\\
        \hspace{\helpSize}($f_7$: \emph{index})
      \end{tabular}

    \end{tabular}

    \\\bottomrule

  \end{tabular}
  \end{figure}

@@

    \begin{tabular}{ll}      
      \begin{tabular}{l}
        \code{[1\, 9]}$\to $\code{2}\\
        \code{[5\, 3\, 8]$\to$3}\\
        \blueCode{$f(\ell) = $($f_5$ $\ell$)}
      \end{tabular}&
      \begin{tabular}{l}
        \code{[[2 1]\, []]}$\to $\code{[2 0]}\\
        \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
        \blueCode{$f(\ell) = $($f_2$ $f_5$ $\ell$)}
      \end{tabular}\\\\

@@

      \begin{tabular}{l}
        \code{0, [9 2 3]$\to $9}\\
        \code{3, [0 2 8 4 5 6]$\to $4}\\
        \blueCode{$f($n$,$l$) = $($f_{10}$ l n)}
      \end{tabular}&
      \begin{tabular}{l}
        \code{1, [9 2 3]$\to $9}\\
        \code{4, [0 2 8 1 5 6]$\to $1}\\
        \blueCode{$f($n$,$l$) = $($f_{10}$ l (+ 1 n))}
      \end{tabular}\\\\
      \begin{tabular}{l}
        \code{3$\to $[-3 -2 -1]}\\
        \code{4$\to $[-4 -3 -2 -1]}\\
        \blueCode{$f($n$) = $($f_{8}$ 0 n)}
      \end{tabular}&
      \begin{tabular}{l}
        \code{2$\to $[2 1]}\\
        \code{4$\to $[4 3 2 1]}\\
        \blueCode{$f($n$) = $($f_{8}$ 0 (- 0 n))}
      \end{tabular}      
    \end{tabular}
    \\\midrule 
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \begin{tabular}{ll}
      \begin{tabular}{l}
        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
          \hspace{\helpSize}($f_0$: \emph{fold})
      \end{tabular}&
      \begin{tabular}{l}
        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
          \hspace{\helpSize}($f_1$: \emph{unfold})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_2$: \emph{map})
      \end{tabular}&
      \begin{tabular}{l}
                  \greenCode{$f_3($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (if (f x) a (cons x a))))}\\
          \hspace{\helpSize}($f_3$: \emph{filter})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
        \hspace{\helpSize}($f_4$: \emph{count upward until predicate holds})
      \end{tabular}&
      \begin{tabular}{l}
        \greenCode{$f_5($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
        \hspace{\helpSize}($f_5$: \emph{length})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($n$,$p$,$l$)\,=\,$($f_1$ p n car l)}\\
        \hspace{\helpSize}($f_6$: \emph{specialization of unfold})
      \end{tabular}&
      \begin{tabular}{l}
        \greenCode{$f_7($p$)\,=\,$($f_4$ ($\lambda$ (x) x) p 0)}\\
        \hspace{\helpSize}($f_7$: \emph{count upward from 0 until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_8($n$,$m$)\,=\,$($f_4$ ($\lambda$ (x) (- n x)) (eq? 0) m)}\\
        \hspace{\helpSize}($f_7$: \emph{count downwards})
      \end{tabular}&
      \begin{tabular}{l}
        \greenCode{$f_9($n$)\,=\,$($f_7$ (eq? n))}\\
        \hspace{\helpSize}($f_9$: \emph{range})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_{10}($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_9$ n) l))}\\
        \hspace{\helpSize}($f_{10}$: \emph{index})
      \end{tabular}&
